---
title: "multiome simulation eda (with null distributions)"
author: "timothy keyes"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Parameters

```{r}
library(tidyverse)
library(tidygraph)
library(patchwork)

source(here::here("R", "utils.R"))
source(here::here("R", "differential_network_utils.R"))

NUM_TFS <- 100
NUM_RES <- 400 # 13000 
NUM_TGS <- 400
NUM_TFS_TO_MODULATE <- 30L
NOISE_SEQUENCE <- seq(0.00, 0.5, 0.05)
SIGNAL_SEQUENCE <- seq(0.00, 0.5, 0.05)

truth <- 
  tibble(
    node_name = paste0("TF", 1:NUM_TFS), 
    is_differential = 
      c(
        rep("yes", NUM_TFS_TO_MODULATE), 
        rep("no", NUM_TFS - NUM_TFS_TO_MODULATE)
      )
  )
```

# Introduction 

In this notebook, we perform identical simulations to those described in `multiome_simulations.Rmd`, but using resampling-based null-distribution estimation to compute a p-value for each TF. These p-values can be used to rank TFs from most to least differential instead of using their raw degree differences (in the separately-constructed networks) or their degree in the edge-subtracted network. 


# First simulation

## create U and V 


```{r}
# create a random value between 0.2 and 0.8 for row of u and v 
# representing the sparsity of that row 
u_sequence <- runif(n = NUM_TFS, min = 0.2, max = 0.8)
v_sequence <- runif(n = NUM_RES, min = 0.2, max = 0.8)

# simulate u
u <- 
  simulate_u(
    num_tfs = NUM_TFS, 
    num_res = NUM_RES, 
    bernoulli_param = u_sequence
  )

# simulate v
v <- 
  simulate_v(
    num_res = NUM_RES, 
    num_tgs = NUM_TGS,
    bernoulli_param = v_sequence
  )
```


## simulate 2 cell types from u and v 

```{r}
u_1b <- 
  u |> 
  add_signal(
    num_tfs_to_modulate = NUM_TFS_TO_MODULATE, 
    prop_edges_to_enrich = 0.2
  ) |> 
  add_noise(addition_prop = 0.1, remove_prop = 0, permutation_prop = 0) |> 
  tidy_u() |> 
  filter(abs(weight) > 0)

u_2b <- 
  u |> 
  add_signal(
    num_tfs_to_modulate = NUM_TFS_TO_MODULATE, 
    prop_edges_to_enrich = 0, 
    prop_edges_to_deplete =  0.2
  ) |> 
  add_noise(addition_prop = 0, remove_prop = 0.2, permutation_prop = 0) |> 
  tidy_u() |> 
  filter(abs(weight) > 0)

# create first baseline cell type by taking the base network and 
# applying random noise
num_edges_to_add <- 
  abs(sum(abs(u) > 0) - nrow(u_1b))

num_edges_to_remove <- 
  abs(sum(abs(u) > 0) - nrow(u_2b))

u_1a <- 
  u |> 
  add_edges(addition_num = num_edges_to_add) |> 
  tidy_u() |> 
  filter(abs(weight) > 0)

u_2a <- 
  u |> 
  remove_edges(remove_num = num_edges_to_remove) |> 
  tidy_u() |> 
  filter(abs(weight) > 0)
```


## Compute null distributions 

```{r}
centrality_enrichment_nulls <- 
  compute_null_distributions(
    weight_matrix = u, 
    num_replicates = 1000, 
    addition_num = num_edges_to_add, 
    remove_prop = 0, 
    permutation_prop = 0, 
    difference_method = "centrality"
  )

edge_enrichment_nulls <- 
  compute_null_distributions(
    weight_matrix = u, 
    num_replicates = 1000, 
    addition_num = num_edges_to_add, 
    remove_prop = 0, 
    permutation_prop = 0, 
    difference_method = "edge"
  )
```

```{r}
centrality_depletion_nulls <- 
  compute_null_distributions(
    weight_matrix = u, 
    num_replicates = 1000, 
    addition_num = 0, 
    remove_num = num_edges_to_remove,
    permutation_prop = 0, 
    difference_method = "centrality"
  )

edge_depletion_nulls <- 
  compute_null_distributions(
    weight_matrix = u, 
    num_replicates = 1000, 
    addition_num = 0, 
    remove_num = num_edges_to_remove,
    permutation_prop = 0, 
    difference_method = "edge"
  )
```


## Compare networks

```{r}
centrality_enrichment_results <- 
  compare_networks(
    edge_list_1 = u_1a, 
    edge_list_2 = u_1b, 
    difference_method = "centrality", 
    null_distributions = centrality_enrichment_nulls
  )

edge_enrichment_results <- 
  compare_networks(
    edge_list_1 = u_1a, 
    edge_list_2 = u_1b, 
    difference_method = "edge", 
    null_distributions = edge_enrichment_nulls
  )

centrality_depletion_results <- 
  compare_networks(
    edge_list_1 = u_2a, 
    edge_list_2 = u_2b, 
    difference_method = "centrality", 
    null_distributions = centrality_depletion_nulls
  )

edge_depletion_results <- 
  compare_networks(
    edge_list_1 = u_2a, 
    edge_list_2 = u_2b, 
    difference_method = "edge", 
    null_distributions = edge_depletion_nulls
  )
```

## Plot results 

### centrality difference

```{r}
centrality_enrichment_results |> 
  plot_roc_curve(pred_col = degree_diff, truth = truth) + 
  labs(subtitle = "raw degree difference")

centrality_enrichment_results |> 
  mutate(score = 1 - p_value_null) |> 
  plot_roc_curve(pred_col = score, truth = truth) + 
  labs(subtitle = "degree difference p-value")

centrality_depletion_results |> 
  plot_roc_curve(pred_col = degree_diff, truth = truth) + 
  labs(subtitle = "raw degree difference")

centrality_depletion_results |> 
  mutate(score = 1 - p_value_null) |> 
  plot_roc_curve(pred_col = score, truth = truth) + 
  labs(subtitle = "degree difference p-value")

```

### edge subtraction 

```{r}
edge_enrichment_results |> 
  plot_roc_curve(pred_col = degree, truth = truth) + 
  labs(subtitle = "raw degree")

edge_enrichment_results |> 
  mutate(score = 1 - p_value_null) |> 
  plot_roc_curve(pred_col = score, truth = truth) + 
  labs(subtitle = "degree p-value")

edge_depletion_results |> 
  plot_roc_curve(pred_col = degree, truth = truth) + 
  labs(subtitle = "raw degree")

edge_depletion_results |> 
  mutate(score = 1 - p_value_null) |> 
  plot_roc_curve(pred_col = score, truth = truth) + 
  labs(subtitle = "degree p-value")
```


```{r}
plot_null_distributions(
  null_distributions = centrality_enrichment_nulls, 
  predictions = mutate(centrality_enrichment_results, degree = abs(degree_diff)),  
  tf_numbers = 
    c(1:15, (nrow(centrality_enrichment_results) - 15):nrow(centrality_enrichment_results))
)
```



```{r}
plot_null_distributions(
  null_distributions = edge_enrichment_nulls, 
  predictions = edge_enrichment_results, 
  tf_numbers = 
    c(1:5, (nrow(edge_enrichment_results) - 5):nrow(edge_enrichment_results))
)
```



# Varying signal 

```{r}

```


# Varying noise 



